Lifting CDCL to Numerical Domains

### 2016-10-09 introduction.tex
contribution

\section{Introduction}
%
A standard static analyses for program safety aims to compute 
invariants by applying transformers that refines an abstract 
value.  Most lattice structure used in static analysis have 
meet and join operations but lack negations.  The meet operation 
precisely model conjunction, and the join operation over-approximate 
disjunction. Thus, the precision loss in static analyzers is often 
due to imprecise join operations. The loss in precision is overcome 
by using richer abstract domains or equipping the analysis with disjunction --
which is often very expensive since the analysis exhibit case enumeration 
behavior.

On the other hand, the combination of precise conjunction in partial assignments 
domain with learning allow Conflict Driven Clause Learning (CDCL)~\cite{cdcl} 
solvers to reason about disjunction without enumerating cases.  
Silva et. al.~\cite{tacas12, sas12, dhk2013-popl} presented 
a formal mechanism for applying abstract interpretation to solve the 
satisfiability problem.  ~\cite{sas12} provide a mathematical generalization 
of CDCL algorithm called, {\em Abstract Conflict Driven Clause Learning} (ACDCL), 
in terms of deduction and abduction transformers, fixed-point, interpolation 
and extrapolation operators on lattices.  The partial assignments in the 
CDCL algorithm are an abstract domain, unit rule is the best abstract 
transformer,  Boolean Constraint Propagation (BCP) computes a greatest fixed-point 
over the abstract domain, decision is dual widening and clause learning can be 
viewed as synthesizing an abstract transformer for negation.  Thus, ACDCL 
lifts learning techniques in SAT solver to operate on non-distributive 
lattice -- thus equipping the analysis with negation. This enables the 
analyzer to refine the analysis and prevent enumeration behavior.

@@

>>>>

\section{Introduction}
%
A standard static analyses~\cite{se2011} for program safety aims to compute 
invariants by applying transformers that refines an abstract 
value.  Most lattice structure used in static analysis have 
meet and join operations but lack negations.  The meet operation 
precisely model conjunction, and the join operation over-approximate 
disjunction. Thus, the precision loss in static analyzers is often 
due to imprecise join operations. The loss in precision is overcome 
by using richer abstract domains or equipping the analysis with disjunction --
which is often very expensive since the analysis exhibit case enumeration 
behavior.

On the other hand, the combination of precise conjunction in partial assignments 
domain with learning allow Conflict Driven Clause Learning (CDCL)~\cite{cdcl} 
solvers to reason about disjunction without enumerating cases.  
Silva et. al.~\cite{tacas12, sas12, dhk2013-popl} presented 
a formal mechanism for applying abstract interpretation~\cite{se2011} to solve the 
satisfiability problem.  ~\cite{sas12} provide a mathematical generalization 
of CDCL algorithm called, {\em Abstract Conflict Driven Clause Learning} (ACDCL), 
in terms of deduction and abduction transformers, fixed-point, interpolation 
and extrapolation operators on lattices.  The partial assignments in the 
CDCL algorithm are an abstract domain, unit rule is the best abstract 
transformer,  Boolean Constraint Propagation (BCP) computes a greatest fixed-point 
over the abstract domain, decision is dual widening and clause learning can be 
viewed as synthesizing an abstract transformer for negation.  Thus, ACDCL 
lifts learning techniques in SAT solver to operate on non-distributive 
lattice -- thus equipping the analysis with negation. This enables the 
analyzer to refine the analysis and prevent enumeration behavior.

@@

Clause learning SAT solvers requires a weak complementation property 
of the partial assignments domain.  However,  the most commonly 
used APRON~\cite{apron} library that implements abstract domains, is specialized 
for abstract interpretation framework which does not require complementation 
operator.  Thus, we implement our own template-based numerical abstract 
domains that support complementation property.  Our proposed technique 
is embodied in our tool called, {\em Titlis}.  To the best of our 
knowledge, this is the first instantiation of CDCL 
algorithm with relational abstract domains.
